---
title: "Manuscript results"
output:
  html_document:
    keep_md: yes
---


```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      out.width = "60%", fig.align = 'center',
                      dpi = 300, dev="jpeg", tidy = TRUE)

library(knitr)
library(tidyverse, quietly = TRUE)
library(lme4)
library(broom.mixed)
library(modelsummary)
library(ggthemes)
library(mgcv)
library(cowplot)

theme_set(theme_classic(base_size = 14))

```


```{r data}

evs <- read_rds("../data/evs-clean.rds")

behavrs_labels <- read_csv("../data/behavior-labels.csv")

# Load mixed effect models estimated in models.R
mixed_3l <- read_rds("../data/mixed-3l-models.rds")
mixed_2l <- read_rds("../data/mixed-2l-models.rds")


```

## Participants

```{r}

sample <- evs %>% 
  select(cntry, id, sex, age) %>% 
  distinct() %>% 
  summarise(n = n(), 
            n_cntry = n_distinct(cntry), 
            female = mean(sex == "Female"), 
            mean_age = mean(age), 
            sd_age = sd(age))
sample
```


We use data on `r sample$n` participants from `r sample$n_cntry` countries in the 1999 wave of the European Values Study, `r sprintf("%.0f", sample$female*100)`% women, mean age `r sprintf("%.1f", sample$mean_age)` with SD `r sprintf("%.1f", sample$sd_age)`.


## Results


```{r tbl1}

behavrs_summary <- evs %>% 
  group_by(behaviour) %>% 
  summarise(common_mean = sprintf("%.2f (%.2f)", 
                             weighted.mean(common, wgt, na.rm = TRUE),
                             Hmisc::wtd.var(common, wgt, na.rm = TRUE)),
            just_mean = sprintf("%.2f (%.2f)", 
                              weighted.mean(just, wgt, na.rm = TRUE), 
                              Hmisc::wtd.var(common, wgt, na.rm = TRUE)),
            common_moral_cor = sprintf("%.2f [%.2f, %.2f]",
                                       cor(common, just, use = "pair"),
                                       cor.test(common, just)$conf.int[1],
                                       cor.test(common, just)$conf.int[2])) 
  
behavrs_summary <- left_join(behavrs_summary, behavrs_labels)

behavrs_summary %>% 
  mutate(group = ifelse(behaviour %in% c("lie", "bribe", "fare_cheat"), 2, 1)) %>% 
  arrange(group, label) %>% 
  select(label, common_mean, just_mean, common_moral_cor) %>% 
  kable(digits = 2,
        col.names = c("Behavior", 
                      "Common M (SD)", 
                      "Moral M (SD)",
                      "Pearson correlation"), 
        caption = "Table 1. Descriptive statistics for perceptions of how common and how moral eight questionable behaviors are, including the correlation (the common-is-moral association).",
        notes = "Note: Based on more than 35,000 respondents in 31 countries in the 1999 wave of the European Values Study. The common-is-moral association is calculated as the Pearson correlation between the two ratings. 95% confidence intervals in brackets.")
  

```

```{r}

evs %>% 
  group_by(behaviour) %>% 
  summarise(r = cor(common, just, use = "pair"),
            just = weighted.mean(just, wgt, na.rm = TRUE)) %>% 
  summarise(min(r), max(r), mean(r), cor(r, just))

```

### The within-individual common-is-moral association is typically positive


```{r}

ind_cor <- evs %>% 
  group_by(cntry, id, relig) %>% 
  summarise(r = cor(common, just), 
            sd_j = sd(just),
            sd_c = sd(common),
            mean_j = mean(just),
            mean_c = mean(common),
            n = n(), 
            .groups = "drop") %>% 
  filter(n > 6) %>% 
  drop_na(r)

ind_cor %>% 
  summarise(mean(r>0), mean(r), sd(r)) %>% 
  kable(digits = 2)

```


```{r fig1, fig.asp=1/3, fig.cap = "Figure 1. Plot showing the average within-individual common-is-moral association in 31 countries in the European Values Study. Each dot represents a country. The box indicates the interquartile range, with the dark line in the box indicating the median."}

cntr_mean_within <- ind_cor %>% 
 group_by(cntry) %>% 
 summarise(mean_r = mean(r, na.rm = TRUE)) 

cntr_mean_within %>% 
  ggplot(aes("Country average", mean_r)) +
  geom_crossbar(stat = "summary", fun.data = "my_summary", width = .4, position = position_nudge(x = .2))+
  gghalves::geom_half_point(side = "l", alpha = .5, transformation = position_identity(), size = 2) +
  geom_hline(yintercept = 0, color = "grey50") +
  labs(x = NULL, y =  "Common-is-moral association") +
  coord_flip() 

```


### The within-individual common-is-moral association is weaker among the more religious.

```{r fig2, fig.cap = "Figure 2. The mean withinbetween-individual common-is-moral association (with 95% bootstrap confidence interval) in each of four subsamples defined by religiosity."}

ind_cor %>% 
  ggplot(aes(relig, r)) +
  stat_summary(fun.data = "mean_cl_boot") +
  labs(x = "Religiosity", 
       y = "Common-is-moral association")

```


```{r fig3, fig.asp = .5, fig.cap = "Figure 3. Fixed effect estimates, with 95% confidence intervals, from the three-level model of moral judgments with Co centered at the individual mean and the rest of the predictors centered at the country means."}

coef_map <- c(
  "common_i" = "Co", 
  "relig" = "Relig",
  "common_i:relig" = "Co x Relig",
  "female" = "Female",
  "age" = "Age",
  "edu_mid" = "Education: Secondary",
  "edu_high" = "Education: Tertiary"
)

modelplot(mixed_3l[[1]], coef_map = coef_map, draw = TRUE) +
  geom_vline(xintercept = 0, color = "grey50") +
  theme_classic(base_size = 14) +
  labs(x = NULL)

# ggsave("3-level-mod-estimates.jpeg", width = 7, height = 3.5, dpi = 300)

```



### The between-individual common-is-moral association is typically positive

```{r}

cm_by_bhvr <- evs %>%
  group_by(cntry) %>% 
  mutate(relig_cntry = mean(relig, na.rm =TRUE)) %>% 
  group_by(cntry, behaviour, relig_cntry) %>%
  summarise(cm = cor(common, just),
            n = n(), 
            se = sqrt((1 - cm^2)/(n - 2)),
            z = cm/se) %>% 
  ungroup()


cm_by_bhvr <- cm_by_bhvr %>% 
  left_join(behavrs_labels) 

cm_by_bhvr %>% 
  group_by(behaviour) %>% 
  summarise(positive = mean(cm > 0)) %>% 
  arrange(positive) %>% 
  mutate(mean(positive))

```


```{r fig4, fig.cap="Figure 4. Boxplots showing how the common-is-moral association for eight questionable behaviors varied across 31 countries in the European Values Study. Boxes indicate the interquartile range, with the dark line in the box indicating the median. Below each box, the exact distribution is represented by dots."}


my_summary <- function(variable) {
  tibble(y = median(variable),
         ymin = quantile(variable, probs = 0.25),
         ymax = quantile(variable, probs = 0.75))
}

cm_by_bhvr %>%  
  mutate(label = fct_rev(label))%>% 
  ggplot(aes(label, cm)) +
  geom_crossbar(stat = "summary", fun.data = "my_summary", 
                width = .4, position = position_nudge(x = .2))+
  gghalves::geom_half_point(side = "l", alpha = .5, 
                            transformation = position_identity(), 
                            size = 2) +
  geom_hline(yintercept = 0, color = "grey50") +
  labs(x = NULL, y =  "Common-is-moral association") +
  coord_flip(ylim = c(min(cm_by_bhvr$cm), max(cm_by_bhvr$cm))) 

# ggsave("cm-boxplot.jpeg", width = 8, height = 4.5, dpi = 350)

```

### The between-individual common-is-moral association is weaker among more religious people

```{r}

cm_by_bhvr_relig <- evs %>%
  drop_na(common, just, relig) %>% 
  group_by(cntry, behaviour, relig) %>%
  summarise(cm = cor(common, just),
            n = n()) %>% 
  ungroup()

est_alpha <- cm_by_bhvr_relig %>%
  ungroup() %>% 
  select(-n) %>% 
  pivot_wider(names_from = behaviour, values_from = cm) %>% 
  select(-cntry, -relig) %>% 
  psych::alpha() %>% 
  .[["total"]] %>% 
  pull(raw_alpha) %>% 
  round(2)

```



```{r fig5,  fig.cap = "Figure 5. The between-individual common-is-moral association for each of eight behaviors in each of four subsamples defined by religiosity, averaged across countries. With 95% bootstrap CI."}

cm_by_bhvr_relig %>% 
  left_join(behavrs_labels) %>% 
  mutate(pl_label = fct_recode(label, 
                            "Driving under influence" = "Driving under influence of alcohol")) %>% 
  ggplot(aes(relig, cm)) +
  stat_summary(fun.data = "mean_cl_boot") +
  facet_wrap(~pl_label, nrow = 2) + 
  geom_hline(yintercept = 0, color = "grey30") +
  labs(x = "Religiosity", y = "Common-is-moral association") + 
  theme_classic() 
#  theme(panel.grid.major = element_line()) 

#ggsave("cm-by-relig.jpeg", width = 8, height = 5.3, dpi = 350)
  
```



```{r fig6, fig.asp = 0.5, fig.cap = "Figure 6. Fixed effect estimates, with 95% confidence intervals, of the common-is-moral association (Co) and its moderation by religiosity (Co x Relig) using mixed-effect model M1. All predictors are centered at the country mean."}

coef_map <- c(
  "common" = "Co", 
  "common:relig" = "Co x Relig"
)

m2l_esimates <- modelplot(set_names(mixed_2l$mod,
                    mixed_2l$behaviour),
          coef_map = rev(coef_map), draw = FALSE) %>% 
  mutate(m = "Base") %>% 
 left_join(behavrs_labels, by = c("model" = "behaviour")) 

panel_lim <- tibble(term = c("Co", "Co", "Co x Relig", "Co x Relig"),
                    estimate = c(-1, 1, -0.3, 0.3)) %>% 
  left_join(m2l_esimates %>% select(term, label))

m2l_esimates %>% 
  ggplot(aes(fct_rev(label), estimate)) +
  geom_pointrange(aes(ymin = conf.low, ymax = conf.high),
                  position = position_dodge(width = .6)) +
  scale_color_grey()+
  coord_flip() +
  geom_hline(yintercept = 0, color = "grey50") +
  # theme_classic(base_size = 11) +
  # theme(panel.grid.major = element_line()) +
  geom_blank(data = panel_lim) +
  facet_wrap(~term, nrow = 1, scales = "free_x") +
  labs(x = NULL, y = NULL)

# ggsave("mod-estimates.jpeg", width = 7, height = 3.5, dpi = 300)

```

```{r fig7, fig.asp=0.53, fig.cap="Figure 7. Religiosity moderates the common-is-moral association. The slope of the blue line represents the estimated common-is-moral association for an average behavior in the average country among non-religious individuals (1 scale unit lower than the average). The red line has less slope, representing the estimated common-is-moral association for an average behavior in the average country among religious individuals (1 scale unit lower than the average)."}

pred_data <- expand_grid(common = seq(-2, 2, .5),
                         relig = c(-1, 1),
                         age = 0,
                         female = 0,
                         edu_mid = 0,
                         edu_high = 0,
                         hdi_gni = 0, 
                         hdi_edu = 0) %>% 
  mutate(common_i = common)

predictions_bw <- mixed_2l %>% 
  transmute(pr = map(mod, 
                   ~mutate(pred_data, 
                           pred = predict(.x, pred_data, 
                                          re.form = ~0)))) %>% 
  unnest(pr) %>% 
  group_by(common, relig) %>%
  summarise(pred = mean(pred)) %>% 
  mutate(effect = "Between-individual")

predictions_wn <- pred_data %>%
  mutate(pred = predict(mixed_3l[[1]], ., re.form = ~0)) %>% 
  mutate(effect = "Within-individual")
  
predictions <- bind_rows(predictions_bw, predictions_wn) %>%
  mutate(relig = factor(relig,
                        labels = c("Non-religious",
                                   "Religious")))
predictions %>%
#  mutate(effect = fct_rev(effect)) %>% 
  ggplot(aes(common, pred, color = relig)) +
  geom_line(size = .7) +
  scale_color_ptol() +
  labs(x = "Common", y ="Moral", color = NULL) +
  theme(legend.position = c(.85, .2)) +
  facet_wrap(~effect)

#ggsave("pred-line.jpeg", width = 7.5, height = 4, dpi = 350)


```



### Country differences in religiosity partly explains country differences in the strength of the common-is-moral association

```{r}

cm_by_bhvr %>% 
  distinct(cntry, relig_cntry) %>% 
  summarise(mean(relig_cntry), sd(relig_cntry)) %>% 
  kable(digits = 2)

```

```{r}

est_alpha <- cm_by_bhvr %>%
  ungroup() %>% 
  select(cntry, behaviour, relig_cntry, cm) %>% 
  pivot_wider(names_from = behaviour, values_from = cm) %>% 
  select(-cntry, -relig_cntry) %>% 
  psych::alpha() %>% 
  .[["total"]] %>% 
  pull(raw_alpha) %>% 
  round(2)

cm_by_relig_cntry = cm_by_bhvr %>% 
  group_by(cntry, relig_cntry) %>% 
  summarise(`Between-individual` = mean(cm)) %>% 
  left_join(cntr_mean_within %>% rename(`Within-individual` = mean_r)) %>% 
  gather("type", "cm", `Between-individual`, `Within-individual`)

est_alpha
cm_by_relig_cntry %>% 
  group_by(type) %>% 
  summarise(tidy(cor.test(relig_cntry, cm))) %>% 
  select(type, estimate, conf.low, conf.high) %>% 
  kable(digits = 2)

```
```{r fig8, fig.cap="Figure 8. Country differences in religiosity partly explain country differences in the strength of the common-is-moral association, whether measured within individuals (left) or between individuals, averaged across behaviors (right)."}

cm_by_relig_cntry %>% 
  mutate(type = fct_rev(type)) %>% 
  ggplot(aes(relig_cntry, cm)) +
  geom_point(alpha = .5) +
  geom_smooth(method = "lm", color = "grey20") +
  facet_wrap(~type) +
  theme_classic() +
  labs(x = "Religiosity", y = "Common-is-moral association")

# ggsave("cntry-relig-effect.jpeg", width = 7.5, height = 4, dpi = 350)

```






## Supplementary tables


```{r stab1, eval=TRUE}

coef_map <- c(
  "common_i" = "Co", 
  "relig" = "Relig",
  "common_i:relig" = "Co x Relig",
  "female" = "Female",
  "age" = "Age",
  "edu_mid" = "Education: Secondary",
  "edu_high" = "Education: Tertiary"
)

# Create df for random effects part
rand_eff_rows <- mixed_3l  %>% 
  map(~tidy(.x, effect = "ran_pars") %>%
        mutate(term = paste0(term, group)) %>% 
        select(term, estimate) %>% 
        filter(!str_detect(term, "cor")) %>% 
        mutate(term = str_replace_all(term,c(
          "sd_+\\(Intercept\\)id:cntry" = "sigma(v1)",
          "sd_+common_iid:cntry" = "sigma(v2)",
          "sd_+\\(Intercept\\)cntry" = "sigma(u1)",
          "sd_+common_icntry" = "sigma(u2)",
          "sd_+ObservationResidual" = "sigma(e)"
        )))) %>% 
  bind_rows(.id = "m") %>% 
    mutate(term = factor(term, levels = unique(term)),
         estimate = sprintf("%.3f",  estimate)) %>% 
  spread(m, estimate, fill =  " ")


fit_rows <-  mixed_3l %>% 
  map(~tibble(term = c("Nk", "Ni", "Nj"),
              estimate = c(nobs(.x), ngrps(.x)))) %>% 
  bind_rows(.id = "m") %>% 
  mutate(term = factor(term, levels = unique(term)),
         estimate = sprintf("%.0f",  estimate)) %>% 
  spread(m, estimate)

rows <- bind_rows(rand_eff_rows, fit_rows)

msummary(mixed_3l, 
         statistic = "conf.int", 
         statistic_vertical = FALSE,
         effect = "fixed", 
         coef_map = coef_map, 
         gof_omit = "IC|Log|REML|sampling|pss|",
         add_rows = rows,
         escape = FALSE,
         output = "kableExtra",
         title = "Supplementary Table 1. Estimates and 95% confidence intervals from different specifications of the three-level models.") %>% 
  kableExtra::kable_classic()


```



```{r stab2, eval=FALSE}

cm_by_bhvr %>% 
  mutate(behaviour = factor(behaviour, levels = behavrs_labels$behaviour)) %>% 
  select(Country = cntry, behaviour, cm) %>% 
  spread(behaviour, cm) %>% 
  select(Country, state_benef:casual_sex) %>% 
  knitr::kable(digits = 2, caption = "Supplementary Table 2. Country estimates of the common-is-moral association for eight questionable behaviors, and the country-mean autonomy index.")


```



```{r stab_mods, results="asis"}

coef_map <- c(
  "common" = "Co", 
  "relig" = "Relig",
  "common:relig" = "Co x Relig",
  "female" = "Female",
  "age" = "Age",
  "edu_mid" = "Education: Secondary", 
  "edu_high" = "Education: Tertiary"
)

order <- c(2, 1, 7, 8, 4, 6, 4, 5)

for(i in order){
mod_list <- list("M1. Baseline" = mixed_2l$mod[[i]], 
                    "M2.1. logit(p_Mo>1)" = mixed_2l$mod_tob_1[[i]], 
                    "M2.2. log(Mo)|Mo > 1" = mixed_2l$mod_tob_2[[i]])

label <- behavrs_labels$label[i]
title <- paste0('Supplementary Table 3. Estimates and 95% confidence intervals from different specifications of the mixed-effects models for "', label, '" behavior.')

# Create df for random effects part
rand_eff_rows <- mod_list  %>% 
  map(~tidy(.x, effect = "ran_pars") %>%
        select(term, estimate) %>% 
        filter(!str_detect(term, "cor")) %>% 
        mutate(term = str_replace_all(term,c(
          "sd_+\\(Intercept\\)(.cntry)?" = "sigma(u1)",
          "sd_+common(.cntry)?" = "sigma(u2)",
          "sd_+Observation(.Residual)?" = "sigma(e)"
        )))) %>% 
  bind_rows(.id = "m") %>% 
    mutate(term = factor(term, levels = unique(term)),
         estimate = sprintf("%.3f",  estimate)) %>% 
  spread(m, estimate, fill =  " ")


fit_rows <-  mod_list  %>% 
  map(~tibble(term = c("Ni", "Nj"),
              estimate = c(nobs(.x), ngrps(.x)))) %>% 
  bind_rows(.id = "m") %>% 
  mutate(term = factor(term, levels = unique(term)),
         estimate = sprintf("%.0f",  estimate)) %>% 
  spread(m, estimate)

rows <- bind_rows(rand_eff_rows, fit_rows)

tb <- msummary(mod_list, 
         statistic = "conf.int", 
         statistic_vertical = FALSE,
         effect = "fixed", 
         coef_map = coef_map, 
         gof_omit = "IC|Log|REML|sampling|pss|",
         add_rows = rows,
         escape = FALSE,
         output = "kableExtra",
         title = title) %>% 
  kableExtra::kable_classic()

cat(tb)
}

```

